---
title: "Good Plot Bad Plot"
author: "Solimar Carrasquillo Ho"
date: "2024-10-23"
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: true
      

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)
```


### Load the libraries
```{r}
library(tidyr)
library(tidyverse)
library(tidytuesdayR)
library(ggplot2)
library(here)
library(ggmap)
library(ggspatial)
library(devtools)
library(dplyr)
library(ggimage)
library(prettydoc)
```


## We're going to be working with _spooOOOoooky_ dataset from the TidyTuesdayR 
```{r}
spooky <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-10-10/haunted_places.csv')
view(spooky)
```

### Lets pull the state data, so we can create a map.
```{r}
state <- map_data("state")
head(state)
```

# Data Manipulation
### We're going to manipulate the data to count the occurances of each spooky place per state.
```{r}
spooky_places <- spooky %>%
  count(state, state_abbrev, sort = TRUE) %>%
  mutate(state = tolower(state))
```

### We have to join the data sets so we can make the data workable! I used an inner join.
```{r}
spooky_states<- state %>%
  #mutate(region = str_to_title(region)) %>% #capitalize the state to match spooky data
  rename(state = region) %>% #rename column to match the spooky_places data
  inner_join(spooky_places) #join the two data sets together

view(spooky_states)

  write.csv(here("Good_plot_bad_plot", "Data", "spooky_states.csv"))
```

# Make your plots!

## Bad Plot: 
Our task was to make Nyssa and everyone cry from how *bad* our ugly plot is....
Ready to cry your eyes out?

```{r}

#I want to make an ugly pie chart that shows the count of spooky places per state. 
bad_plot <- ggplot(spooky_states, aes(x = "", y = n, fill = state)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y") +
  labs(title = "Haunted Places by State") +
 theme(
    plot.title = element_text(color = "orange", hjust = 0.5),
    axis.title.y = element_text(color = "white"),
    axis.title.x = element_text(color = "white") # Set title color to white
  )
super_ugly <- ggbackground(bad_plot, background ="../Data/81118061-xl-72dpi_360x.png")
super_ugly

ggsave(filename = here("Good_plot_bad_plot", "Output", "super_ugly.png"))
```

### Why would we never make a plot like this ever, _ever again?_
* We absolutely hate pie charts. There are never exceptions for creating a pie chart. 
* The image background is absolutely horrific. Not because of the amount of patriotism, but because it makes the data completely unreadable. We love patriotism....**GO VOTE!!!!!**
* The axis labels are white, which make it readable, but we can't see the other values along the pie chart.
* Each state has its own color, which makes it confusing to read. 
* The count of spooky places per state has been log transformed and that makes it unclear which state has the highest count of spooky places.

Now that we're crying and know why we're crying. Let's move onto our _good and beautiful_ plot. :)

## Good Plot: 

```{r}
#I'm making a median value so the color gradient looks better! :)
median_value <- median(spooky_states$n)

#Let's make a map and look at the count of spooky places in each state.
ggplot()+
  geom_polygon(data = spooky_states,
            aes(x = long,
                y = lat,
                group = group,
                fill = n), color = "black") +
  coord_map() + 
  theme_void() +
  scale_fill_gradient2(mid = 'orange',
                       low = 'darkred',
                       high = "black", midpoint = median_value, limits=c(-25, 1070)) +
  theme(plot.title = element_text(color = "orange", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  labs(title = "Haunted Places Across the U.S.", 
    subtitle = "A Spooky Overview",
    fill = "Haunted Places Count"
  )

#Save the plot now
ggsave(filename = here("Good_plot_bad_plot", "Output", "super_beautiful.png"))

```

### Why is this plot so good?
* This plot has clear labels!
* Legend includes a color gradient, which shows which state has the highest count of spooky places.
* Image is clear and easy to decipher. You can tell that California is the spookiest state in the U.S.


