---
title: "Tidy Tuesday 2024-11-05"
author: "Solimar Carrasquillo Ho"
date: "2024-11-10"
output: html_document
---

Tidy tuesday rubric:

dataset
export knitted file
plot and export 
place proper code in folder
describe what was learned in that week's tidy tuesday.

# Libraries
```{r}
library(here)
library(tidyverse)
library(janitor) #to clean the data 
library(ggplot2)
library(stringr)
library(dplyr)
library(knitr) #use to make a table
library(kableExtra)
library(patchwork)
```


# Load the data. We're going to be looking at the _Democracy and Dictatorship_ dataset from TidyTuesday Nov. 5th 2024.
```{r}
democracy_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-05/democracy_data.csv')
View(democracy_data)

```

I'm curious to know the progression of female leadership over these democracies. 
```{r}
democracy_df <- democracy_data %>%
  select("country_name", "year", "regime_category", "is_monarchy", "monarch_name", "monarch_accession_year", "is_democracy", "is_presidential", "president_name", "president_accesion_year", "is_female_monarch", "is_female_president") 

print(democracy_df)
```

# Data mining and cleaning 

Should I try making a table to summarize all the female leaders? I haven't practiced creating tables since we've learned about them in class. 
We're going to rename the columns too.

```{r}

# Group by leader and country, and count the number of years served
female_monarch <- democracy_df %>%
  filter(is_female_monarch == TRUE) %>%
  group_by(monarch_name, country_name) %>%
  summarise(total_years_served = n(), .groups = "drop")

  colnames(female_monarch) <- colnames(female_monarch) %>%
  str_replace_all("_", " ") %>%  # Replace underscores with spaces
  str_to_title()     

# View the resulting dataset with total years served for each leader
print(female_monarch)


```
Female presidents 
```{r}
female_president <- democracy_df %>%
  filter(is_female_president == TRUE) %>%
  group_by(president_name, country_name) %>%
  summarise(total_years_served = n(), .groups = "drop")

colnames(female_president) <- colnames(female_president) %>%
  str_replace_all("_", " ") %>%  # Replace underscores with spaces
  str_to_title()     

# View the resulting dataset with total years served for each leader
print(female_president)
```




Lets make a table of Monarch using the knitr package.
```{r}
kable(female_monarch, format = "html", caption = "Female Monarch")
```

Now for the presidents!
```{r}
kable(female_president, format = "html", caption = "Female President")
```
I'm having problems creating a plot because I removed the spaces from the column titles, so I'm going to add them back in so I can make my plot!

```{r}
female_president_df <- democracy_df %>%
  filter(is_female_president == TRUE) %>%
  group_by(president_name, country_name) %>%
  summarise(total_years_served = n(), .groups = "drop") %>%
  arrange(desc(total_years_served)) %>%  # Sort by total years served in descending order
  distinct(president_name, .keep_all = TRUE) %>%  # Remove duplicates, keeping only the first occurrence of each president
  slice_head(n = 10)

female_monarch_df <- democracy_df %>%
  filter(is_female_monarch == TRUE) %>%
  group_by(monarch_name, country_name) %>%
  summarise(total_years_served = n(), .groups = "drop") %>%
  arrange(desc(total_years_served)) %>%  # Sort by total years served in descending order
  distinct(monarch_name, .keep_all = TRUE) %>%  # Remove duplicates, keeping only the first occurrence of each monarch
  slice_head(n = 10)

```



We're going to try to learn how to make a lollipop plot this week.
I try and create two lollipop plots to show the length of each female leader. 


Add country name in the bubble at the end!

```{r}

# Combine monarch and country names for better labeling
female_monarch_clean <- female_monarch_df %>%
  mutate(monarch_country = paste(monarch_name, "(", country_name, ")", sep = ""))


# Create the lollipop plot without country name next to leader name
plot1 <- ggplot(female_monarch_clean, aes(x = reorder(monarch_name, total_years_served), y = total_years_served)) +
  geom_segment(aes(x = monarch_name, xend = monarch_name, y = 0, yend = total_years_served), color = "skyblue") +
  geom_point(color = "darkblue", size = 4) +
  geom_text(aes(label = country_name), vjust = -0.5, hjust = 0.5, color = "black", size = 3) +  # Add country name inside bubble
  coord_flip() +  # Flip coordinates for readability
  labs(
    title = "Years of Service by Female Monarchs", 
    caption = "Source: Democracy and Dictatorship Dataset TidyTuesday 2024-11-05",
    x = "Monarch Name",  # Display only the monarch's name
    y = "Total Years Served"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15),
    axis.title = element_text(size = 12)
  )

```

Let's do it for female presidents! 

```{r}
# Combine president and country names for better labeling
female_president_clean <- female_president_df %>%
  mutate(president_country = paste(president_name, "(", country_name, ")", sep = ""))

# Create the lollipop plot
plot1 <- ggplot(female_president_clean, aes(x = reorder(president_name, total_years_served), y = total_years_served)) +
  geom_segment(aes(x = president_name, xend = president_name, y = 0, yend = total_years_served), color = "skyblue") +
  geom_point(color = "darkblue", size = 4) +
  geom_text(aes(label = country_name), vjust = -0.5, hjust = 0.5, color = "black", size = 3) +  # Add country name inside bubble
  coord_flip() +  # Flip coordinates for readability
  labs(
    title = "Years of Service by Female presidents", 
    caption = "Source: Democracy and Dictatorship Dataset TidyTuesday 2024-11-05",
    x = "president Name",  # Display only the president's name
    y = "Total Years Served"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15),
    axis.title = element_text(size = 12)
  )

print(plot2)
```
Now let's use the patchwork to put the two plots next to each other :).

```{r}
plot1 / plot2
```
Finally, we're going to save the plot so we can actually see them! 

```{r}
ggsave(here("Tidy_Tuesday", "Output", "female_leaders.png"))
```


For this week, I learned how to make tables (again) and how to utilize lollipop plots. I think I should have filtered the data more to show less countries. It could be more informative. 