---
title: "Week_07_classwork"
author: "Solimar Carrasquillo Ho"
date: "2024-10-08"
output:
    prettydoc::html_pretty:
    theme: tactile
    highlight: github
  
---

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
library(ggdogs)
library(prettydoc)
```

# Load the data
###Don't forget to copy over the data from the Fall 2024 files to your personal Data folder! :)
#If you don't *the data won't load!*
```{r}
popdata<-read_csv(here("Week_07","Data","CApopdata.csv"))
```
```{r}
stars<-read_csv(here("Week_07","Data","stars.csv"))
```

# view data 
```{r}
world <-map_data("world")
head(world)
```

```{r}
usa <-map_data("usa")
head(usa)
```

#States
```{r}
states<-map_data("state")
```

#county
```{r}
county<-map_data("county")
```

#Italy
```{r}
italy <-map_data("italy")
head(italy)
```

#Make a map of the world 
```{r}
ggplot()+
  geom_polygon(data = world, aes(x = long, y = lat, group = group))
```

#what happens if you forget the group? 

```{r}
ggplot() +
  geom_polygon(data = world, aes(x = long, y = lat))
```
# Let's add color! 
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, y = lat, group = group, fill = region),
               color = "black") +
  guides(fill = FALSE) + 
  theme(panel.background = element_rect(fill = "lightblue"))

```
#Change the map projection (there are many types of projection)
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, y = lat, group = group, fill = region),
               color = "black") +
  guides(fill = FALSE) + 
  theme(panel.background = element_rect(fill = "lightblue"))+
coord_map(projection = "mercator",
            xlim = c(-180,180))

```
#Use the sinusoidal projection

```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long,
                   y = lat,
                   group = group, 
                   fill = region),
               color = "black")+
  theme_minimal()+
  guides(fill = FALSE)+
  theme(panel.background = element_rect(fill = "lightblue"))+
  coord_map(projection = "sinusoidal",
            xlim = c(-180,180))

```
#Lets make a map of California for location of sea stars
```{r}
head(states)
```

#filter out just California data
```{r}
CA_dates <- states %>%
  filter(region == "california")
```

#lets make a simple map of CA
```{r}
ggplot()+ 
  geom_polygon(data = CA_dates, aes(x = long, y = lat, group = group)) +
  coord_map()
```

#Were going to add multiple layers by _wrangling the data_
##this created a new data frame 

```{r}
CApop_county <- popdata %>%
  select("subregion" = County, Population) %>%
  inner_join(county) %>%
  filter(region == "california")
head(CApop_county)
```
#Make a map of CA population by county
##We'll log transform the data so we can visualize the populations better!
```{r}
ggplot()+
  geom_polygon(data = CApop_county,
               aes(x = long,
                   y= lat,
                   group = group,
                   fill = Population),
               color = "black")+
  coord_map()+
  theme_void()+ 
    scale_fill_gradient(trans = "log10")
```


#Now lets layer starfish data onto this map!
## We need to visiualize the data 
```{r}
head(stars)
```

# We'll add on the star data set

```{r}
ggplot()+
  geom_polygon(data = CApop_county,
               aes(x = long,
                   y= lat,
                   group = group,
                   fill = Population),
               color = "black")+
  geom_point(data = stars,
             aes(x = long,
                 y = lat,
                 size = star_no)) +
  coord_map()+
  theme_void()+ 
    scale_fill_gradient(trans = "log10")
```

#Now lets make the plot a little more beautiful 

```{r}
ggplot()+
  geom_polygon(data = CApop_county,
               aes(x = long,
                   y= lat,
                   group = group,
                   fill = Population),
               color = "black")+
  geom_point(data = stars,
             aes(x = long,
                 y = lat,
                 size = star_no)) +
  coord_map()+
  theme_void()+ 
    scale_fill_gradient(trans = "log10")+
  labs(size = "# stars/m2") 
  ggsave(here("Week_07","Output","CApop.pdf"))
  
```

# AWESOME DOG PACKAGE 
```{r}
ggplot(mtcars) +
  geom_dog(aes(mpg, wt), dog = "eyes", size = 7)
```

