---
title: "Week_07_Homework Map Practice"
author: "Solimar Carrasquillo Ho"
date: "2024-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries

```{r}
library(tidyverse)
library(here)
library(devtools)
library(ggmap)
library(ggspatial)
```

## Let's load the data. Apparently there are tons of bigfoot sightings...which I'm skeptical about so let's explore this dataset! 
```{r}
bigfoot <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-13/bigfoot.csv')
view(bigfoot)

```

## We're going to pull the state data from the dataset
```{r}
state <-map_data("state")
head(state)
```



### Data Manipulation
#### We're going to look at the seasons and how many bigfoot sightings over time
##### From this data, I can see the amount of sightings by state. It would be interesting to plot the frequency on a map based on state!I should look into add the lat and long
```{r}
bigfoot <- bigfoot %>% 
  group_by(state, season) %>% #group state and seasons
  filter(season== "Summer" | season == "Winter"  | season == "Spring") %>% #removes unknown season data
  summarise(sighting= n()) #create a new column with sightings 

bigfoot_s <- state %>%
  mutate(region = str_to_title(region)) %>% #capitalize the state to match bigfoot data
  rename(state = region) %>% #rename column to match the bigfoot data
  inner_join(bigfoot) #join the two data sets together
view(bigfoot_s)
  
```

# Perrrfect! Now that we've manipulated the data set in a way we can use it...let's make a map! 
#I want to look at the bigfoot sightings across the US
```{r}
ggplot()+
  geom_polygon(data = bigfoot_s,
               aes(x = long, y = lat, group = group,
                                     fill = sighting), #should fill the states based on number of sightings
               color = "black") #color of the outline of the states
```

#Hmm, I want to make the points larger from the sightings...
```{r}
ggplot()+
  geom_polygon(data = bigfoot_s, aes(x = long, y = lat, group = group, fill = sighting), #should fill the states based on number of sightings
               color = "black") +
  coord_map(projection = "mercator") +
  facet_wrap(~ season)+ #mapping by seasons
  labs(fill = "Bigfoot Sightings",
       title = "Seasonal Bigfoot Sightings in the USA") +
  theme(plot.title = element_text(hjust = 1, vjust = 1),
        legend.title = element_text(size =7))
```

## Based on our **super kewl** map, we can see that bigfoot sightings are more prominent in the west coast during the summertime.
### This is probably because bigfoot are spending more time outside...hunting and gathering for whatever it is they eat hahah. Super _legit_. 
 