---
title: "Week_07_classwork_2"
author: "Solimar Carrasquillo Ho"
date: "2024-10-14"
output:
    prettydoc::html_pretty:
    theme: tactile
    highlight: github
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries

```{r}
library(tidyverse)
library(here)
library(prettydoc)
library(devtools)
library(ggmap)
library(ggspatial)
library(emojifont)
```


# Load that data girl ! :-)
## Use the plastic pollution data for actual homework

```{r}
ChemData <-read_csv(here("Week_07", "Data", "chemicaldata_maunalua.csv"))
view(ChemData)
```


## lets make a map of Oahu (only going to pull out the coordinate system)
```{r}
Oahu <-get_map("Oahu")
```
#make a map
```{r}
ggmap(Oahu)
```

# Lets get the coordinates to focus on Wailupe 
## The "zoom" argument goes from 3 to 20. 3 = continent level; 20 = single building level
```{r}
WP <- data.frame(long = -157.7621, lat = 21.27427)
Map1 <- get_map(WP, zoom = 17)
```

#change the map type so it's prettier

```{r}
Map1<- get_map(WP, zoom = 17, maptype = "satellite")
ggmap(Map1)
```
#now lets make a stamen water color map 
## why didn't this work?

```{r}
Map1 <- get_map(WP, zoom = 17, maptype = "stamen_watercolor", source = "stadia")
```

#Lets plot salinity values of each sampling points

```{r}
Map1 <- get_map(WP, zoom = 17, maptype = "satellite")
ggmap(Map1)+
  geom_point(data = ChemData,
             aes(x = Long,
                 y = Lat, 
                 color = Salinity),
             size = 4) +
  scale_color_viridis_c()
```

# Lets add a scale bar and north arrow
# bl = bottom left 
# tl = top left 
# crs = 4326 is code for a
```{r}
ggmap(Map1)+
  geom_point(data = ChemData,
             aes(x = Long,
                 y = Lat, 
                 color = Salinity),
             size = 4) +
  scale_color_viridis_c() +
  annotation_scale(bar_cols = c ("yellow", "white"), 
                    location = "bl") + 
  annotation_north_arrow(location = "tl")+ 
  coord_sf(crs = 4326)
```

# You can use geocode() if you don't know the exact GPS coordinates

```{r}
geocode("the white house")
```

#UH Manoa
```{r}
geocode("University of Hawaii at Manoa")
```

#another day, another R package that slays

```{r}
search_emoji("smile")
ggplot() +
  geom_emoji('smiley', 
             x = 1:5, 
             y = 1:5,
             size = 11)
```

