---
title: "Week_13_class"
author: "Solimar Carrasquillo Ho"
date: "2024-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set up the scripts
```{r}
library(tidyverse)
library(here)
```
This week we're learning about for loops! :)
We'll start with an easy for loop, which tells us what year it is.
```{r}
print(paste("The year", 2000))
```

```{r}
years<-c(2015:2021)

for (i in years){ #set up the for loop where i is the index
  print(paste("The year is", i))
  }
```
We made it print, but it's not saving it anywhere. Let's make it save.
```{r}
#create space for the for loop with an empty matrix that is as long as the year vector
year_data <-tibble(year = rep(NA, length(years)), #column name for year
                   year_name = rep(NA, length(years))) #column name for the year 

year_data
```

lets create a line by line...
```{r}
for (i in 1:length(years)){ #set up the for loop where i is the index
  year_data$year_name[i]<-paste("The year is", years[i])
}

year_data
```

How can you use a for loop in your file? 
We can write a for loop when reading in multiple csv files. This can be helpful for my independent project! 

```{r}
testdata <- read_csv(here("Week_13", "Data", "cond_data", "011521_CT316_1pcal.csv"))

glimpse(testdata)
```
#now we're going to tell r where each of the files are! 
1. figure out where our path is with the here function
2. 

```{r}
CondPath <- here("Week_13", "Data", "cond_data")

# list all the files in that path with a specific patter
# in this case were looking for everything with .csv in the filename!

#we'll use regrex to be more specific
files <- dir(path = CondPath, pattern = ".csv")
files
```

# were going to calculate the mean temperature and salinity from each file and save it! 

```{r}
# pre allocate space
# make an empty space
cond_data<- tibble(filename = rep(NA, length(files)),
                   mean_temp = rep(NA, length(files)),
                   mean_sal = rep(NA, length(files)),
                   )

cond_data
```

# were going to write a basic code to calculate the mean and build it out. Were going to plug in 1 to test this file. 

```{r}
raw_data<-read_csv(paste0(CondPath,"/", files[3]))
head(raw_data)
```

```{r}
mean_temp<- mean(raw_data$Temperature, na.rm = TRUE)
mean_temp
```
Now were going to turn it into a for loop! 
```{r}
for (i in 1:length(files)){
  raw_data<-read_csv(paste0(CondPath,"/", files[i]))
glimpse(raw_data)
  
} #loop over 1:3 the number of files

```
# add in the columns now

```{r}
for (i in 1:length(files)){
  raw_data<-read_csv(paste0(CondPath,"/", files[i]))
glimpse(raw_data)

cond_data$filename[i]<- files[i]
  
} #loop over 1:3 the number of files

cond_data
```

now were going to add in the means

```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
#glimpse(raw_data)
cond_data$filename[i]<-files[i]
cond_data$mean_temp[i]<-mean(raw_data$Temperature, na.rm =TRUE)
cond_data$mean_sal[i]<-mean(raw_data$Salinity, na.rm =TRUE)
} 
cond_data
```


# We can use the purrr package that will help you write a for loop, but it's w/ the tidyverse style language and in less lines of code! 

Were going to calculate a mean from random set numbers and do it 10 times

create a vector
```{r}
1:10 
```
then make it random based on a normal distribution. it also uses the pipe tidy. with this you don't have to preallocate space either!
```{r}
1:10 %>%
  map(rnorm, n =15)

```

# now lets calculate the mean from each list randomly!!!!
```{r}
1:10 %>%
  map(rnorm, n = 15) %>%
  map_dbl(mean) # calculate the mean, and it's now a vector whic is the type double
```

now were going to make our own function....
```{r}
1:10 %>% # list 1:10
  map(function(x) rnorm(15, x)) %>%
  map_dbl(mean)
```

Were going to use a formula so we want to change the arguements w/in the function
```{r}
1:10 %>%
  map(~ rnorm(15, .x)) %>% # changes the arguments inside the function
  map_dbl(mean)
```
lets bring in the files using purrr instead of a forloop
```{r}
CondPath<-here("Week_13", "Data", "cond_data")
files <- dir(path = CondPath, pattern = ".csv")

files
```

So now we can get the full file names in one less step by doing thiss......
```{r}
files <- dir(path = CondPath,pattern = ".csv", full.names = TRUE)
#save the entire path name
files
```

next were going to read in the files using map instead of the for loop
```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") # map everything to a dataframe and put the id in a column called filename

data
```

# calculate means
```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") %>% # map everything to a dataframe and put the id in a column called filename
  group_by(filename) %>%
  summarise(mean_temp = mean(Temperature, na.rm = TRUE),
            mean_sal = mean(Salinity,na.rm = TRUE))
data
```

homework will show the 