---
title: "Week_13_hmwk"
author: "Solimar Carrasquillo Ho"
date: "2024-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load your libraries
```{r}
library(tidyverse)
library(here)
```

The final homework assignment!!! 
we have to take the 4 csv files in data/homework and calculate the mean and standard deviation of both temperatures and light for EACH tidepool. 
Use BOTH a for loop and map() functions in your script. So do this twice!! 

```{r}
tidepool <- read_csv(here("Week_13", "Data", "homework", "TP1.csv"))

glimpse(tidepool)
```


```{r}
TP <- here("Week_13", "Data", "homework")

# list all the files in that path with a specific patter
# in this case were looking for everything with .csv in the filename!

#we'll use regrex to be more specific
files <- dir(path = TP, pattern = ".csv")
files
```

# were going to calculate the mean & standard deviation for temperature and light intensity from each file and save it! 
Temp.C and Intensity.lux
```{r}
# pre allocate space
# make an empty space
tide_data<- tibble(filename = rep(NA, length(files)),
                   mean_temp = rep(NA, length(files)),
                   mean_light = rep(NA, length(files)),
                   sd_temp = rep(NA, length(files)),
                   sd_light = rep(NA, length(files))
                   )

tide_data
```

# were going to write a basic code to calculate the mean and SD and build it out. Were going to plug in 1 to test this file. 

```{r}
raw_data<-read_csv(paste0(TP,"/", files[1]))
head(raw_data)
```

Now lets write the simple calculations for the mean and SD 
```{r}
mean_temp<- mean(raw_data$Temp.C, na.rm = TRUE)
mean_temp

sd_temp <- sd(raw_data$Temp.C, na.rm = TRUE)
sd_temp

mean_light <-mean(raw_data$Intensity.lux, na.rm = TRUE)
mean_light

sd_light <-sd(raw_data$Intensity.lux, na.rm = TRUE)
sd_light
```

Now were going to turn it into a for loop! 
```{r}
for (i in 1:length(files)){
  raw_data<-read_csv(paste0(TP,"/", files[i]))
glimpse(raw_data)
  
} #loop over 1:4 the number of files

```


now lets add in the data
```{r}
for (i in 1:length(files)){
  raw_data<-read_csv(paste0(TP,"/", files[i]))
glimpse(raw_data)

tide_data$filename[i]<- files[i]
  
} #loop over 1:3 the number of files

tide_data
```

```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(TP,"/",files[i]))
#glimpse(raw_data)
tide_data$filename[i]<-files[i]
tide_data$mean_temp[i]<-mean(raw_data$Temp.C, na.rm =TRUE)
tide_data$mean_light[i]<-mean(raw_data$Intensity.lux, na.rm =TRUE)
tide_data$sd_temp[i]<-sd(raw_data$Temp.C, na.rm =TRUE)
tide_data$sd_light[i]<-sd(raw_data$Intensity.lux, na.rm =TRUE)
} 
tide_data
```

Now let's move onto using the maps function to calculate the mean and sd. 

lets bring in the files using purrr instead of a forloop
```{r}
TP<-here("Week_13", "Data", "homework")
files <- dir(path = TP, pattern = ".csv")

files
```
So now we can get the full file names in one less step by doing thiss......
```{r}
files <- dir(path = TP,pattern = ".csv", full.names = TRUE)
#save the entire path name
files
```

next were going to read in the files using map instead of the for loop
```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") # map everything to a dataframe and put the id in a column called filename

data
```

# calculate means and sd for Temp.C and Intensity.lux
```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") %>% # map everything to a dataframe and put the id in a column called filename
  group_by(filename) %>%
  summarise(mean_temp = mean(Temp.C, na.rm = TRUE),
            mean_light = mean(Intensity.lux, na.rm = TRUE),
            sd_temp = sd(Temp.C, na.rm = TRUE),
            sd_light = sd(Intensity.lux, na.rm = TRUE))
data
```