---
title: "Week_13"
author: "Solimar Carrasquillo Ho"
date: "2024-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Intro to modeling 

# load up the libraries 
```{r}
library(modelsummary) # to show model output
library(tidymodels) # for tidy models
library(broom) # for clean model output
library(flextable) # to look at model results in a nice table
library(performance) # to check model assumptions
library(here)
library(pandoc) # asked to download to run line 76
library(wesanderson)
```
The basic anatomy of a linear model
1. To run a simple linear model for the following formula
mod <-lm(y~x, data = df)
lm = linear model, y = dependent variable, x = independent variables(s), df = dataframe
You'll read this as y is a function of x. 

Multiple regression
mod <- lm(y~x1 + x2, data = df)

Interaction term
mod <-lm(y~x*x2, data = df) 

Let's try to model the penguin dataset.

```{r}
Peng_mod<-lm(bill_length_mm ~ bill_depth_mm*species, data = penguins)
```


We need to check the assumptions of the model with performance! ALWAYS check the assumptions of your specific models. 

```{r}
check_model(Peng_mod) #checks the assumption
```

View results in base r
```{r}
anova(Peng_mod)
```
Use summary to look at the coeffcients (effect size) with error
```{r}
summary(Peng_mod)
```

The broom package allows us to take the summary, residual, and ANOVA to put it in a nice tibble. Which allows you to do something with it in the data and export your results!

```{r}
coeffs<-tidy(Peng_mod) #just put tidy() around the model
coeffs
```
If we wnat to look at the pvalue for the entire model and r values, we'll use the glance model
```{r}
results <-glance(Peng_mod)
results
```

If we want to extract the residuals for the fitted values. We'll use the augment function
```{r}
resid_fitted<-augment(Peng_mod)
resid_fitted
```

Results in the modelsummary package. 

This will export an actual document in the format of a table!!! 
```{r}
Peng_mod_noX<- lm(bill_length_mm ~ bill_depth_mm, data = penguins)

# make a list of models and name them
models <-list("Model with interactions" = Peng_mod,
              "MOdel with no interaction" = Peng_mod_noX)

#save the results as a .docx
modelsummary(models, output = here("Week_13", "Output", "table.docx"))
```

We can make a canned coefficient modelplot.

```{r}
modelplot(models) +
    labs(x = 'Coefficients', 
         y = 'Term names') +
    scale_color_manual(values = wes_palette('Darjeeling1'))

```
What if we want to run models with many parameters. 

```{r}
models <- penguins %>%
  ungroup() %>% #the penguin data are grouped so we'll need to ungroup them
  nest(.by = species) # nest all the data by species

models
```

```{r}
models <- penguins %>%
  ungroup() %>% #the penguin data are grouped so we'll need to ungroup them
  nest(.by = species) %>% # nest all the data by species
  mutate(fit = map(data, ~lm(bill_length_mm~body_mass_g, data = .)))

models
```

```{r}
models$fit
```
to get the results
```{r}
results<-models %>%
  mutate(coeffs = map(fit, tidy), #look at the coefficients
         modelresults = map(fit, glance)) %>% # R2 and other values
  select(species, coeffs, modelresults) %>%
           unnest()

view(results)
```

```{r}
lm_mod<-linear_reg() %>%
  set_engine("lm") %>%
  fit(bill_length_mm ~ bill_depth_mm*species, data = penguins)

lm_mod
```

