---
title: "Week_09_Homework"
author: "Solimar Carrasquillo Ho"
date: "2024-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use any dataset you would like and make two useful functions. One that is a plot and one that is not. Turn in your functions using RMarkdown with at least 2 examples per function showing its flexibility. Due, as always, at 1pm on Tuesday.

## Homework tasks 
you have to make two useful functions 

## Load the libraries
```{r}
library(ggplot2)
library(devtools)
library(dplyr)
library(tidyr)
library(tidyverse)
```

## Load the dataset
```{r}
cheeses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-06-04/cheeses.csv')
view(cheeses)
```

## Data Manipulation
### I want to look at the cheese types, but noticed there are multiple types of animals in the milk column. We'll need to separate this to make the data usable! 

Separate all rows from data set. 

```{r}
all_cheese <- cheeses %>%
  separate_rows(country, sep = ", ") %>%
  separate_rows(milk, sep = ", ")
```

The first function will clean the data and create a count of how many cheeses per country. 
1. first function to pull names of cheese and country
2. second function to pull from first function to get animal milk
```{r}
cheese_cleaned <- function(data, specific_country = NULL){ data %>%
  filter(is.null(specific_country)| country == specific_country) %>% #Filter for specific country
  group_by(country)
}
```

Pull the France cheeses! 

```{r}
France_cheese <- cheese_cleaned(cheeses, "France")
```



What about creating a function to find out more about cheeses for those w/ dietary preferences?

```{r}
filter_cheeses <- function(data, milk = NULL, country = NULL, vegan = NULL, vegetarian = NULL) {
  
  # Start with the full dataset
  filtered_data <- data
  
  # Apply milk type filter if specified
  if (!is.null(milk)) {
    filtered_data <- filtered_data[filtered_data$milk == milk, ]
  }
  
  # Apply country filter if specified
  if (!is.null(country)) {
    filtered_data <- filtered_data[filtered_data$country == country, ]
  }
  
  # Apply vegan filter if specified
  if (!is.null(vegan)) {
    filtered_data <- filtered_data[filtered_data$vegan == vegan, ]
  }
  
  # Apply vegetarian filter if specified
  if (!is.null(vegetarian)) {
    filtered_data <- filtered_data[filtered_data$vegetarian == vegetarian, ]
  }
  
  # Return the filtered dataset
  return(filtered_data)
}


```

Now let's start using that filter! 
```{r}
french_vegetarian_cow_cheeses <- filter_cheeses(data = cheeses, country = "France", milk = "cow", vegetarian = TRUE)

print(french_vegetarian_cow_cheeses)

```



# Create the function that will make a plot!
```{r}
plot_cheeses_by_milk_and_country <- function(data) {
  #Count cheeses by milk type and country
  milk_counts <- data %>%
    count(country, milk, name = "count")
  
  #Make a plot
  ggplot(milk_counts, aes(x = country, y = count, color = milk)) +
    geom_point(size = 3) +
    labs(
      title = "Number of Cheeses by Milk Type and Country",
      x = "Country",
      y = "Number of Cheeses"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for readability
    scale_color_brewer(palette = "Set1")  # Optional: Use a color palette for milk types
}

# Example usage:
# plot_cheeses_by_milk_and_country(cheeses)


```


Now lets make the plot! 
```{r}
plot_cheeses_by_milk_and_country(all_cheese)

```

This is a little messy...so let's add a filter that will only look at the countries that have more than 30 types of cheeses....

```{r}
plot_cheeses_by_milk_and_country <- function(data) {
  # Step 1: Count total cheeses by country
  country_counts <- table(data$country)
  
  # Step 2: Filter countries with more than 20 cheeses
  countries_with_more_than_30 <- names(country_counts[country_counts > 30])
  
  # Step 3: Count cheeses by milk type and country for filtered countries
  milk_counts <- as.data.frame(table(data$country[data$country %in% countries_with_more_than_30],
                                      data$milk[data$country %in% countries_with_more_than_30]))
  
  colnames(milk_counts) <- c("country", "milk", "count")
  
  # Step 4: Create a scatter plot
  ggplot(milk_counts, aes(x = country, y = count, color = milk)) +
    geom_point(size = 3) +
    labs(
      title = "Number of Cheeses by Milk Type and Country (Countries with >30 Cheeses)",
      x = "Country",
      y = "Number of Cheeses"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for readability
    scale_color_brewer(palette = "Set1")  # Optional: Use a color palette for milk types
}
```

run it!

```{r}
plot_cheeses_by_milk_and_country(all_cheese)

```

