---
title: "Week_09_homework"
author: "Solimar Carrasquillo Ho"
date: "2024-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Homework is due on 11/2! 

Will have to work with words to clean up issues with character data to reorder factors into orders that make sense. 

column names could be better named. 

* make a plot where one of the axes is a factor
* use either intertidal.csv r intetidal_latitude.csv (include data dictionary)
* if choosing to visualize anything by tide height, it should go from low to mid to high tide. 

use stringr package to clean names 

# load the libraries
```{r}
library(tidyverse)
library(stringr)
library(ggplot2)
library(here)
library(janitor) #to clean the data
library(tidytext)
```



# load the data
```{r}
intertidal_lat <- read_csv(here("Week_09", "Data","intertidaldata_latitude.csv"))
head(intertidal_lat)
intertidal <- read_csv(here("Week_09", "Data","intertidaldata.csv"))
head(intertidal)
```

# Data Cleaning!

Some columns include (count), and quadrat columns have extra numbers or spaces. 
```{r}
clean_intertidal <- intertidal %>%
  clean_names() %>%
  mutate(quadrat = str_remove_all(quadrat, "[\\.0-9\\s]")) #line will remove unwanted characters 
view(clean_intertidal)
```

Cute! We have to make the axis a factor, so we'll pick the quadrats! We only want to look at a select part of the data so we can continue manipulating here...

```{r}
quadrat <- clean_intertidal %>%
  pivot_longer(cols = c(bare_rock: anemone), #put all organisms into one column
               names_to = "Cover",
               values_to = "Percent") %>% 
  filter(site == "Bodega") %>% #lets only look at one location
  select(c(site, transect, quadrat, Cover, Percent)) %>%
  mutate(quadrat = factor(quadrat)) %>% # we need to convert this into a factor :)
  droplevels()  #always add droplevels

#we're adding this line in to reorder the axis lable!
quadrat$quadrat <- factor(quadrat$quadrat, levels = c("Low", "Mid", "High"))

  
```

Sweeeet! Now we can really make the plot. 
```{r}
ggplot(quadrat,
       aes(x = Cover,
           y = Percent,
           fill = quadrat)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Intertidal Organism Coverage at Bodega",
       x = "Intertidal Organisms",
       y = "Total Percent Cover (%)",
       fill = "Tide") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(
   labels = c("algae" = "Algae",
               "anemone" = "Anemone",
               "bare_rock" = "Bare Rock",
               "gooseneck_barnacles" = "Gooseneck Barnacles",
               "large_barnacles" = "Large Barnacles",
               "mussels" = "Mussels",
               "small_barnacles" = "Small Barnacles"))
```

