---
title: "TT_5"
author: "Solimar Carrasquillo Ho"
date: "2024-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#loading the libraries
```{r}
library(ggplot2)
library(showtext) #easier way to use new fonts
library(tidyr)
library(dplyr)
library(here)
```


# Load in the data! 

```{r}
cbp_resp <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-26/cbp_resp.csv')
View(cbp_resp)
```

# If I have time I want to learn to use google fonts
I've loaded in the package showtext, which we can use the following webpage to search available texts 
https://fonts.google.com/

```{r}
font_add_google("Bungee", "bungee")
showtext_auto()
showtext_opts(dpi=300)

body_font <- "bungee"
```


# data cleaning. Let's look at immigration from central america  

```{r}
data <- cbp_resp %>%
  select(fiscal_year, citizenship, encounter_type, demographic, title_of_authority, encounter_count)
View(data)
```

Filter out our central american countries.
going to do just mexico and venezula! 
```{r}
central_america <- data %>%
    filter(citizenship %in% c("MEXICO", "VENEZUELA")) %>% #"COLOMBIA",  "HONDURAS", "GUATEMALA", "EL SALVADOR")) %>%
  filter(encounter_count != 0) %>%   #remove if any of the encounter_count was 0 
  mutate(citizenship = str_to_title(citizenship)) #remove all caps

view(central_america)
```


# lets calculate the total encounter_count for each country
```{r}
count <- central_america %>% 
  group_by(fiscal_year, citizenship, encounter_type) %>%
  summarise(
    total_encounters = sum(encounter_count), .groups = 'drop'
  ) %>%
  mutate(
    lower_bound = total_encounters * 0.9, # 10% below total_encounters
    upper_bound = total_encounters * 1.1  # 10% above total_encounters
  )



view(count)
```




# lets make an easy plot
```{r}
map <- ggplot(count, aes(x = fiscal_year, y = total_encounters)) +
  geom_line(aes(color = citizenship), size = 1) +  # Line plot for values
  geom_point(aes(color = citizenship), size = 2) + # Points on the line
  facet_wrap(~ citizenship, ncol = 2) +                      # Facet by country
  labs(
    title = "Fiscal Year Trends by Country",
    x = "Fiscal Year",
    y = "Total Encounters",
    color = "Country"
  ) +
  theme_minimal() +                            # Use a clean theme
  theme(
    strip.text = element_text(size = 3, face = "bold"), # Customize facet labels
    axis.text.x = element_text(angle = 45, hjust = 1)   # Rotate x-axis labels
  )

map
```


# Now I want to try learning how to make a ribbon plot. 
# In order to make a ribbon plot, you'll need to create an lower and upper bound...
```{r}
ggplot(count, aes(x = fiscal_year, y = total_encounters, group = encounter_type)) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = encounter_type), alpha = 0.3) + # Ribbon
  geom_line(aes(color = encounter_type), size = 1) +  # Line plot for mean total_encounters
  geom_point(aes(color = encounter_type), size = 2) + # Points for total_encounters
  labs(
    title = "Immigration Encounters Over Time by Citizenship",
    x = "Fiscal Year",
    y = "Total Encounters",
    fill = "Encounter Type",
    color = "Encounter Type"
  ) +
  facet_wrap(~ citizenship, ncol = 2) +  # Facet wrap by citizenship with 3 columns
  theme_minimal() + # Clean theme
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    strip.text = element_text(size = 12, face = "bold") # Adjust facet label size
  )

```
# trying to remove the log
```{r}
ggplot(count, aes(x = fiscal_year, y = total_encounters, group = encounter_type)) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = encounter_type), alpha = 0.3) + # Ribbon
  geom_line(aes(color = encounter_type), size = 1) +  # Line plot
  geom_point(aes(color = encounter_type), size = 2) + # Points
  labs(
    title = "Immigration Encounters Over Time by Citizenship",
    x = "Fiscal Year",
    y = "Total Encounters",
    fill = "Encounter Type",
    color = "Encounter Type"
  ) +
  facet_wrap(~ citizenship, ncol = 3) +  # Facet by citizenship with 3 columns
  scale_y_continuous(labels = scales::comma) +  # Ensure linear scale with comma-separated values
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    strip.text = element_text(size = 12, face = "bold")
  )

```
```{r}
plot <- ggplot(count, aes(x = fiscal_year, y = total_encounters, group = encounter_type)) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = encounter_type), alpha = 0.3) + # Ribbon
  geom_line(aes(color = encounter_type), size = 1) +  # Line plot for mean total_encounters
  geom_point(aes(color = encounter_type), size = 2) + # Points for total_encounters
  labs(
    title = "Immigration Encounters Over Time by Citizenship",
    x = "Fiscal Year",
    y = "Total Encounters",
    fill = "Encounter Type",
    color = "Encounter Type"
  ) +
  facet_wrap(~ citizenship, ncol = 3) +  # Facet wrap by citizenship
  scale_y_continuous(labels = scales::comma) +  # Linear scale for y-axis
  theme_minimal() +
  theme(
    plot.title = element_text(size = 11),  # Reduced title size by 3 points
    axis.title.x = element_text(size = 11), # Reduced x-axis label size by 3 points
    axis.title.y = element_text(size = 11), # Reduced y-axis label size by 3 points
    axis.text.x = element_text(size = 8),   # Reduced x-axis text size by 3 points
    axis.text.y = element_text(size = 8),   # Reduced y-axis text size by 3 points
    legend.position = "right",  # Move legend to the right
    legend.box = "vertical",   # Stack the legend vertically
    legend.text = element_text(size = 7),  # Reduced legend text size by 3 points
    legend.title = element_text(size = 8), # Reduced legend title size by 3 points
    strip.text = element_text(size = 9)    # Reduced facet label size by 3 points
  )


```
So we learned how to use google fonts and how to create a ribbon plot today!

Now let's save the plot...
```{r}
ggsave(here("Tidy_Tuesday", "TT_5", "Output", "immigration_encounters_plot.png"), plot = plot, width = 10, height = 8)

```

