---
title: "Tidy Tuesday 2024-10-29"
author: "Solimar Carrasquillo Ho"
date: "2024-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the libraries
```{r}
library(ggplot2)
library(here)
library(dplyr)
library(tidyr)
library(tidyverse)
library(gganimate)
library(gapminder)
```

# Load the haunted movie data

```{r}
monster_movies <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-29/monster_movies.csv')
View(monster_movies)
```

# Data manipulation 
Lets look at the last 30 years worth of scary movies and clean up the titles 
separate by , in genres
```{r}
clean_monsters <- monster_movies %>%
  drop_na() %>%
  separate_rows(genres, sep = ",\\s*") %>%
  filter(year > 1994,
         year < 2025)
```

I want to try using the gganimate package to show how genres of monster movies changed over the 30-year period! 

```{r}
ggplot(clean_monsters, 
       aes(year, average_rating)) +
  geom_point() +
  scale_color_manual(values = year) +
  facet_wrap(~ genres) 

```
Let's narrow down some of the genres to focus on.

```{r}
scary_monsters <- clean_monsters %>%
  filter(genres %in% c("Action", "Crime", "Horror", "Mystery", "Sci-Fi", "Thriller"))
```

Ok! Let's try make the facet wrap again, but add the animation! 
I had to add {round(frame_time)} to remove the years in decimal points.

```{r}
ggplot(scary_monsters, 
       aes(x = year, y = average_rating, fill = genres)) +  
  geom_point(size = 3, alpha = 0.7) +                    
  facet_wrap(~ genres) +
  labs(title = 'Movie Ratings Over Time By Genre: Year {round(frame_time)}', 
       subtitle = 'Years 1995 - 2024',
       x = 'Year',
       y = 'Average Rating',
      caption = 'Data provided by Monster Movie 10-29-2024 TidyTuesday') +  
  scale_fill_manual(values = c(Action = "blue", 
                                "Crime" = "red", 
                                "Horror" = "purple", 
                                "Mystery" = "green", 
                                "Sci-Fi" = "orange", 
                                "Thriller" = "brown")) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),  # Center title
    plot.subtitle = element_text(size = 14, hjust = 0.5),              # Center subtitle
    plot.caption = element_text(size = 10, hjust = 0.5),               # Center caption
    strip.text = element_text(face = "bold"),
    legend.position = "none"  # Remove the legend
  ) +
  transition_time(year) +
  ease_aes('linear')

```


Now lets save our plot!
```{r}
ggsave(here("Tidy_Tuesday", "Output", "20241029_monstermovies.png"), plot = last_plot(), height = 8, width = 7)

```

