---
title: "Week_08 Classwork"
author: "Solimar Carrasquillo Ho"
date: "2024-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### We're going to be learning how to do text analysis *yay!*
## Let's install the following packages. 
```{r cars}
library(here)
library(tidyverse)
library(tidytext)
library(wordcloud2)
library(janeaustenr)
library(ggirl)
library(patchwork)
library(ggrepel)
library(gganimate)
library(magick)
library(palmerpenguins)
library(sourrr)
```

### What is string?
A string is the same as a character. You can tell its a string by the use of quotations.
```{r}
words <- "This is a string"
words
```
# You can have several strings into a vector

```{r}
words_vector <- c("Apples", "Bananas", "Oranges")
words_vector
```
##### There are *four basic families of functions* in the stringr package.

* _Manipulation_ which allow you to manipulate individual characters within the strings in character vectors.
* _Whitespace tools_ to add, remove, and manipulate whitespace.
* _Locale sensitive operations_ whose operations will vary from locale to locale.
* _Pattern matching functions_ These recognize four engines of pattern description. The most common is regular expressions, but there are three other tools.

Let's look at the manipulation function first. . .
```{r}
paste ("High temp", "Low pH")
```
Add a dash in between the words

```{r}
paste ("High temp", "Low pH", sep = "-")

```

You can also remove spaces in between words by using a zero (0)

```{r}
paste0("High temp", "Low pH")

```
We can make the manipulate function to add in data such as...
```{r}
shapes <- c("Square", "Circle", "Triangle")
paste("My favorite shape is a", shapes)
```

It can be useful when making labels for your plots... how can we use this in our labels though.
```{r}
two_cities <- c("best", "worst")
paste("It was the", two_cities, "of times.")
```
If you want to know how long a string is

```{r}
str_length(shapes)
```
What if you want to extract specific characters 
```{r}
seq_data <-c("ATCCCGTC")
str_sub(seq_data, start = 2, end = 4) #this extracts the 2nd to 4th
```
You can also modify the strings
?This didn't work...
```{r}
str_sub(seq_data, start = 3, end = 3) <- "A"
seq_data
```
You can duplicate the patterns in the strings. Lets duplicate it 2 & 3 times
```{r}
str_dup(seq_data, times = c(2, 3)) #number of times to duplicate
```

Lets pretend you have some words with extra white spaces and R thinks its an entirely new word...let's see how we would deal with that
```{r}
badtreatments <-c("High", " High", "High ", "Low", "Low")
badtreatments
```
Removing the white spaces...
```{r}
str_trim(badtreatments)
```

You can remove from one side or the other
```{r}
str_trim(badtreatments, side = "left") # this removes left
```
The opposite of str_trim is str_pad
```{r}
str_pad(badtreatments, 5, side = "right") # add a white space to the right side after the 5th character

```

Adding characters instead of white space
```{r}
str_pad(badtreatments, 5, side = "right", pad = "1") # add a 1 to the right side after the 5th character

```

Make everything uppercase
```{r}
x <- "I love R!"
str_to_upper(x)
```
Lowercase
```{r}
str_to_lower(x)
```

capitlaize the first letter of each word
```{r}
str_to_title(x)
```
Now let's do some pattern matching using _stringr_ function. 
This will allow us to view, detect, locate, extract, match, replace, ans split strings...
```{r}
data<- c("AAA", "TATA", "CTAG", "GCTT")
```

```{r}
str_view(data, pattern = "A")
```
Detect a specific pattern
```{r}
str_detect(data, pattern = "A")
```

Locate a pattern
```{r}
str_locate(data, pattern = "AT")
```

### Regex: regular expressions
* Metacharacters - match a single character (. \ ( ) { $ * + ? })
* Sequences
* Quantifiers
* Character Classes
* POSIX character classes 

Lets say you have the following set of strings..
```{r}
vals <-c("a.b", "b.c", "c.d")
```

We want to replace the "." with a space! This will only replace the first "."
```{r}
str_replace(vals, "\\.", " ")
```
If there's multiple "." in the string then we have to use antoher function
```{r}
vals<-c("a.b.c", "b.c.d","c.d.e")
#string, pattern, replace
str_replace(vals, "\\.", " ")
```

There are shorthand versions (called anchors) for commonly used sequences in R. You'll need to look them up! 

Now lets subset the vector to only keep strings with digits..
```{r}
val2 <-c("test 123", "test 456", "test")
str_subset(val2, "\\d")
```
A _character class_ are a list of characters enclosed by square brackets []. They're used to only match one of the different characters. 

```{r}
strings<-c("550-153-7578",
         "banana",
         "435.114.7586",
         "home: 672-442-6739")
```

Lets make a regex that finds strings and contain a phone number..

```{r}
phone <- "([2-9][0-9]{2})[-.](0-9]{3})[-.]([0-9]{4})"
```
```{r}
str_detect(strings, phone)

```
let's subset strings w/ phone numbers
```{r}
test <- str_subset(strings, phone)
test
```
think pair share...
replace all "." with "-" and extract only the numbers. remove any white space! 
```{r}
test %>%
  str_replace_all(pattern = "\\.", replacement= "-") %>%
  str_replace_all(pattern = "[a-zA-Z]|\\:", replacement = "") %>% #removes things we don't want
  str_trim() #trim the white space
```
Lets explore books by Jane Austen now

```{r}
head(austen_books())
tail(austen_books())
```
clean it up
```{r}
original_books <- austen_books() %>%
  group_by(book) %>%
  mutate(line = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                 ignore_case = TRUE)))) %>%
  ungroup()

head(original_books)
```
We'll use the function unnest_tokens(), which will unnest the data so its only one word per row. 
```{r}
tidy_books <- original_books %>%
  unnest_tokens(output = word, input = text) #add a column named word, with the input column as text 

head(tidy_books)
```

We'll use the function get_stopwords() to remove common, but unnecessary words such as and, by, therefore...
```{r}
head(get_stopwords())
```

We can now join to remove all the stopwords
```{r}
cleaned_books <- tidy_books %>%
  anti_join(get_stopwords())

head(cleaned_books)
```

Lets find the most common words across all the jane austen books

```{r}
cleaned_books %>% 
  count(word, sort = TRUE)
```

How do we modify this code to count the most popular words by book? What about by each chapter w/in a book? 

Lets try using the function get_sentiments() for a sentiment analysis! 
```{r}
sent_word_counts <- tidy_books %>%
  inner_join(get_sentiments()) %>% #this will only keep pos or negative words
  count(word, sentiment, sort = TRUE) #will count the words

head(sent_word_counts)[1:3,]
```

Now let's plot it!


```{r}
sent_word_counts %>%
  filter(n > 150) %>% #take only if there are over 150 instances
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>% #add a column where it'll count if word is negative
  mutate(word = reorder(word, n)) %>% #sort largest to smallest
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col() +
  coord_flip() +
  labs(y = "Contribution to sentiment")
```
Now make a word cloud!

```{r}
words <-cleaned_books %>%
  count(word) %>% 
  arrange(desc(n)) %>% #sort the words
  slice(1:100) #lets only look at top 100

wordcloud2(words, shape = 'triangle', size = 0.3)
```

## Totally awesome R Package is ggirl which will turn your plot into a postcard!!


### Moving onto the online class video

_Advanced plotting packages_
* patchwork - bringing plots together
* ggrepel - repel your labels on your plots
* gganimate - add some animations to plots
* magick - photoprocessing 

Let's start with two simple plots from palmer penguin dataset.

Plot 1
```{r}
p1<-penguins %>%
  ggplot(aes(x = body_mass_g, 
             y = bill_length_mm, 
             color = species))+
  geom_point()
p1
```


Plot 2
```{r}
p2<-penguins %>%
  ggplot(aes(x = sex, 
             y = body_mass_g, 
             color = species))+
  geom_jitter(width = 0.2)
p2
```
Lets add them together. 

```{r}
p1 + p2
```

```{r}
p1 / p2 +
 plot_layout(guides = 'collect') + #makes only one legend
  plot_annotation(tag_levels = 'A') #adds table letters
```

We really only want one legend.. 
patchwork will allow you to process the 

```{r}
p1+p2 +
  plot_layout(guides = 'collect') + #makes only one legend
  plot_annotation(tag_levels = 'A') #adds table letters
```

## ggrepel 
```{r}
view(mtcars)
```

Make a simple plot
```{r}
ggplot(mtcars, aes(x = wt,
                   y = mpg,
                   label = rownames(mtcars))) + #how you extract rownames
  geom_text() + #creates a text label
  geom_point(color = 'red')
```

Lets use the ggrepel package to repel the labels from the points so we can acutally see the labels! yay.

```{r}
ggplot(mtcars, aes(x = wt,
                   y = mpg,
                   label = rownames(mtcars))) + #how you extract rownames
  geom_text_repel() + #repel the labels
  geom_point(color = 'red')
```

You can also make them into labels!
```{r}
ggplot(mtcars, aes(x = wt,
                   y = mpg,
                   label = rownames(mtcars))) + #how you extract rownames
  geom_label_repel() + #repel the labels
  geom_point(color = 'red')
```

## gganimate

Lets show how this changes over the years! 
```{r}
penguins %>%
  ggplot(aes(x = body_mass_g,
             y = bill_depth_mm,
             color = species)) +
  geom_point() +
  transition_states(
    year, #what are we animating by
    transition_length = 2, #the relative length to go from one to the next in seconds
    state_length = 1 #the length of how long itll stay at the point
  )
```

Now lets change the movement of the animation.
```{r}
penguins %>%
  ggplot(aes(x = body_mass_g,
             y = bill_depth_mm,
             color = species)) +
  geom_point() +
  transition_states(
    year, #what are we animating by
    transition_length = 2, #the relative length to go from one to the next in seconds
    state_length = 1 #the length of how long itll stay at the point
  ) +
  ease_aes("bounce-in-out") 
```

Now finally lets add a label to show the year over time. 
```{r}
penguins %>%
  ggplot(aes(x = body_mass_g,
             y = bill_depth_mm,
             color = species)) +
  geom_point() +
  transition_states(
    year, #what are we animating by
    transition_length = 2, #the relative length to go from one to the next in seconds
    state_length = 1 #the length of how long itll stay at the point
  ) +
  ease_aes("bounce-in-out") +
  labs(title = 'Year: {closest_state}') #this is the arguement of state of where you are...Year is the static part...
```

And now to save! 
```{r}
penguins %>%
  ggplot(aes(x = body_mass_g,
             y = bill_depth_mm,
             color = species)) +
  geom_point() +
  transition_states(
    year, #what are we animating by
    transition_length = 2, #the relative length to go from one to the next in seconds
    state_length = 1 #the length of how long itll stay at the point
  ) +
  ease_aes("bounce-in-out") +
  labs(title = 'Year: {closest_state}') #this is the arguement of state of where you are...Year is the static part...
anim_save(here("Week_08", "Output", "mypenguingif.gif"))
```

Last package magick
You can put images on top of your plots! 
google an image and the link has to end in png

```{r}
penguin<-image_read("https://pngimg.com/uploads/penguin/pinguin_PNG9.png")
penguin
```
First we have to save the plot we're going to use before adding the image.
```{r}
penguins %>%
  ggplot(aes(x = body_mass_g,
             y = bill_depth_mm,
             color = species)) +
  geom_point()

ggsave(here("Week_08", "Output", "penguinplot.png"))
```
Now, that's its saved as a png file. We can add the image w/ the magick package.

```{r}
penplot <-image_read(here("Week_08", "Output", "penguinplot.png"))
out <- image_composite(penplot, penguin, offset = "+40+30")
out
```
We can also add GIFs to plot 
```{r}
pengif <-image_read("https://media3.giphy.com/media/H4uE6w9G1uK4M/giphy.gif")
outgif <-image_composite(penplot, pengif, gravity = "center") #gravity as where does it pull. center puts it right in the center
animation <- image_animate(outgif, fps = 10, optimize = TRUE)
animation
```

## totally awesome R package ! 
## SOURDOUGH RECIPE MIXTURE OMGGGG YESSS
```{r}
build_recipe(final_weight = 1000, hydration = 0.75)
```

